Racial and Ethnic Differences in ETI: Outcomes

### 2024-10-24 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
stratified by age group

```{r setup}
#| include: false
# Libraries
library(tidyverse)
library(gt)
library(gtsummary)
library(ggflowchart)
library(lmerTest)
library(splines)
library(ggeffects)
# Paths
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/tim/CF"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

The following models are adjusted for age group and insurance status. We fit a Gaussian mixed model for continuous outcomes with random intercept for each subject.

```{r results='asis'}
#| message: false
invisible(lapply(continuous_outcomes[1:3], function(o) {
  # Formula
  f <-
    as.formula(paste0(
      o,
      "~ bs(Days, knots = 0, degree = 1) * Race + age_group + Insurance + (1 | eDWID)"
    ))
  m <- lmer(f, data = encounter)
  # Print table
  cat("\n")
  cat("\n")
  cat("##", o)
  cat("\n")
  cat("\n")
  cat("### Model results")
  cat("\n")
  cat("\n")
  m %>%
    tbl_regression() %>%
    print()
  cat("\n")
  cat("\n")
  # Plot
  cat("\n")
  cat("\n")
  cat("## Model predictions")
  cat("\n")
  cat("\n")
  pred <- predict_response(m, terms = c("Days[-1095,0,730]", "Race"))
  pred %>% print()
  cat("\n")
  cat("\n")
  cat("### Plot")
  cat("\n")
  cat("\n")
  pred %>%
    plot() %>%
    print()
  cat("\n")
  cat("\n")
}))
```

>>>>

```{r setup}
#| include: false
# Libraries
library(Hmisc)
library(tidyverse)
library(gt)
library(gtsummary)
library(ggflowchart)
library(lmerTest)
library(splines)
library(ggeffects)
# Paths
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/timvigers/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

```{r}
# Create change in slope variable
encounter$slope_change_days <- (encounter$Days > 0) * encounter$Days
label(encounter$slope_change_days) <- "Change in Slope"
encounter$Years <- encounter$Days / 365.25
encounter$slope_change_years <- (encounter$Years > 0) * encounter$Years
label(encounter$slope_change_years) <- "Change in Slope"
```

The following models are adjusted for age group and insurance status. We fit a Gaussian mixed model for continuous outcomes with random intercept for each subject (random slope did not improve the model AIC).

## FEV1 % predicted

### Full cohort

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1_ppred_rn ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = gli_fev1_ppred_rn, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1_pct_predicted ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = GLI_FEV1_pct_predicted, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

### Age < 18 at ETI initiation

The following models were not adjusted for age group.

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race +
    Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = gli_fev1_ppred_rn, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1_pct_predicted ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = GLI_FEV1_pct_predicted, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

### Age >= 18 at ETI initiation

The following models were not adjusted for age group.

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1_ppred_rn ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = gli_fev1_ppred_rn, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1_pct_predicted ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = GLI_FEV1_pct_predicted, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

## FVC % predicted

### Full cohort

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fvc_ppred_rn ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = gli_fvc_ppred_rn, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FVC_pct_predicted ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = GLI_FVC_pct_predicted, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

### Age < 18 at ETI initiation

The following models were not adjusted for age group.

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fvc_ppred_rn ~ Years * Race + slope_change_years * Race +
    Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = gli_fvc_ppred_rn, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FVC_pct_predicted ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = GLI_FVC_pct_predicted, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

### Age >= 18 at ETI initiation

The following models were not adjusted for age group.

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fvc_ppred_rn ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = gli_fvc_ppred_rn, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FVC_pct_predicted ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = GLI_FVC_pct_predicted, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

## FEV1/FVC % predicted

### Full cohort

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1fvc_ppred_rn ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = gli_fev1fvc_ppred_rn, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1FVC_pct_predicted ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = GLI_FEV1FVC_pct_predicted, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

### Age < 18 at ETI initiation

The following models were not adjusted for age group.

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1fvc_ppred_rn ~ Years * Race + slope_change_years * Race +
    Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = gli_fev1fvc_ppred_rn, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1FVC_pct_predicted ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = GLI_FEV1FVC_pct_predicted, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

### Age >= 18 at ETI initiation

The following models were not adjusted for age group.

#### GLI 2022 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1fvc_ppred_rn ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = gli_fev1fvc_ppred_rn, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

#### GLI 2012 equations

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1FVC_pct_predicted ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = GLI_FEV1FVC_pct_predicted, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

## BMI %ile

### Full cohort

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  bmipercentile ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = bmipercentile, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

### Age < 18 at ETI initiation

The following models were not adjusted for age group.

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  bmipercentile ~ Years * Race + slope_change_years * Race +
    Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = bmipercentile, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

### Age >= 18 at ETI initiation

The following models were not adjusted for age group.

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  bmipercentile ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = bmipercentile, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

## Height %ile

### Full cohort

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  heightpercentile ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = heightpercentile, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

### Age < 18 at ETI initiation

The following models were not adjusted for age group.

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  heightpercentile ~ Years * Race + slope_change_years * Race +
    Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = heightpercentile, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

### Age >= 18 at ETI initiation

The following models were not adjusted for age group.

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  heightpercentile ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = heightpercentile, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

## Weight %ile

### Full cohort

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  weightpercentile ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = weightpercentile, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

### Age < 18 at ETI initiation

The following models were not adjusted for age group.

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  weightpercentile ~ Years * Race + slope_change_years * Race +
    Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = weightpercentile, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

### Age >= 18 at ETI initiation

The following models were not adjusted for age group.

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  weightpercentile ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = weightpercentile, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```
